---
title: "ANLY 512 - Problem Set 1"
subtitle: "Anscombe's quartet"
author: "Sonam Gupta"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Rmarkdown


```{r}
summary(cars)
```

You can also embed plots, for example:

```{r, echo=FALSE, fig.width=4, fig.height=3}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Questions

1. Anscombes quartet is a set of 4 $x,y$ data sets that were published by *Francis Anscombe* in a 1973 paper [*Graphs in statistical analysis*](https://www.jstor.org/stable/2682899?seq=1#page_scan_tab_contents).  For this first question load the `anscombe` data that is part of the `library(datasets)` in `R`.  And assign that data to a new object called `data`.

```{r}
library(datasets)
data <- anscombe
```

2. Summarise the data by calculating the mean, variance, for each column and the correlation between each pair (eg. x1 and y1, x2 and y2, etc) (Hint: use the `fBasics()` package!)
```{r}
dataDf <- data.frame(data)
library(fBasics)
# mean of every column
colAvgs(dataDf) 
# variance of every column
colVars(dataDf)

#colMeans(dataDf) #gives means for each column in the dataframe created above #one more way to do the same task 

#correlations between pairs:
cor(dataDf$x1, dataDf$y1)
cor(dataDf$x2, dataDf$y2)
cor(dataDf$x3, dataDf$y3)
cor(dataDf$x4, dataDf$y4)
```

3. Create scatter plots for each $x, y$ pair of data.
```{r}
plot(dataDf$x1, dataDf$y1)
plot(dataDf$x2, dataDf$y2)
plot(dataDf$x3, dataDf$y3)
plot(dataDf$x4, dataDf$y4)
```

4. Now change the symbols on the scatter plots to solid circles and plot them together as a 4 panel graphic
```{r}
#Using ggplot2 package since it allows to create multiple plots at the same time
library(ggplot2)
library(gridExtra)
plot1 <-  ggplot(dataDf) + geom_point(aes(x1, y1))
plot2 <-  ggplot(dataDf) + geom_point(aes(x2, y2))
plot3 <-  ggplot(dataDf) + geom_point(aes(x3, y3))
plot4 <-  ggplot(dataDf) + geom_point(aes(x4, y4))
grid.arrange(plot1, plot2, plot3, plot4, top = "Anscombe's Scatter PLots")
```

5. Now fit a linear model to each data set using the `lm()` function.
```{r}
lm(formula = y1 ~ x1, dataDf)
lm(formula = y2 ~ x2, dataDf)
lm(formula = y3 ~ x3, dataDf)
lm(formula = y4 ~ x4, dataDf)
```
6. Now combine the last two tasks.  Create a four panel scatter plot matrix that has both the data points and the regression lines. (hint: the model objects will carry over chunks!)
```{r}
plot1 <- ggplot(dataDf) + geom_point(aes(x1, y1)) + geom_abline(intercept = 3, slope = 0.5)
plot1
plot2 <- ggplot(dataDf) + geom_point(aes(x2, y2)) + geom_abline(intercept = 3, slope = 0.5)
plot2
plot3 <- ggplot(dataDf) + geom_point(aes(x3, y3)) + geom_abline(intercept = 3, slope = 0.49)
plot3
plot4 <- ggplot(dataDf) + geom_point(aes(x4, y4)) + geom_abline(intercept = 3, slope = 0.49)
plot4
```

7.  Now compare the model fits for each model object.
```{r, results="asis"}
library("stargazer")
stargazer(dataDf, type = "text")

```

8. In text, summarize the lesson of Anscombe's Quartet and what it says about the value of data visualization.

To summarize what I learned from Anscombes Quartet is that,reading from the statistics table does not always give the complete ore desired information about the statistics of the data. Visualizations on the other hand provides you with a graphical representation of what your data is like and what all features it has. Outlier detection becomes really easy with plots and other visualizations. 







