---
title: "Quantified Self"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: cerulean
---

```{r setup, include=FALSE}
library(flexdashboard)
library(ggplot2)
library(plotly)
# Creating an empty dataframe
newSelfDf <- data.frame(cupsOfCoffee = integer(0), hoursSleep = integer(0), hoursInSocialMedia = integer(0), hoursOnResearch = numeric(0), minsInNovelReading = numeric(0), workout = integer(0))

# loop to create rows
for(i in 1:365)
{
	# get cupsOfCoffee
	cupsOfCoffee = round(rnorm(1, 2, 0.5))
	
	# workout yes or no
	if(runif(1) < 0.9)
	{
		workout = 1
	}
	else
	{
		workout = 0
	}
	
	
	# get hoursSleep
	if ( runif(1) < 0.85)
	{
		meanHrs = 7.5
		stdHrs = 0.75
		if (cupsOfCoffee >= 3)
		{
			meanHrs = meanHrs - runif(1, min = 1, max = 1.5)
		}
		else if (cupsOfCoffee == 2)
		{
			meanHrs = meanHrs - runif(1, min = 0.75, max = 1)
		}
		else
		{
			stdHrs = 0.25
		}
		
		if (workout == 0)
		{
			meanHrs = meanHrs - runif(1, min = 0.25, max = 1)
		}
		hoursSleep = round(rnorm(1, meanHrs, stdHrs) * 2) / 2
	}
	else
	{
		meanHrs = 4.5
		stdHrs = 0.5
		if (cupsOfCoffee >= 3)
		{
			meanHrs = meanHrs - runif(1, min = 1, max = 1.5)
		}
		else if (cupsOfCoffee == 2)
		{
			meanHrs = meanHrs - runif(1, min = 0.75, max = 1)
		}
		else
		{
			stdHrs = 0.25
		}
		hoursSleep = round(rnorm(1, meanHrs, stdHrs) * 2) / 2
	}
	
	# get hoursInSocialMedia, research, novel reading
	if (workout == 1)
	{
		minSocialMedia = 1
		maxSocialMedia = 2
		minResearch = 5
		maxResearch = 9
	}
	else
	{
		minSocialMedia = 2
		maxSocialMedia = 4
		minResearch = 3
		maxResearch = 7
	}
	
	hoursInSocialMedia = sample(minSocialMedia:maxSocialMedia, 1, replace = TRUE)
	hoursOnResearch = sample(minResearch:maxResearch, 1, replace = TRUE)
	
	if (hoursOnResearch >= 6 )
	{
		minNovel = 15 
		maxNovel = 60
	}
	else
	{
		minNovel = 45
		maxNovel = 120
	}
	minsInNovelReading =  ceiling(    sample(minNovel:maxNovel, 1, replace = TRUE)     /15)*15    
		
	rowsDataset <- c(cupsOfCoffee, hoursSleep, hoursInSocialMedia, hoursOnResearch, minsInNovelReading, workout)
	newSelfDf <- rbind(newSelfDf, rowsDataset)
}

names(newSelfDf) <- c("cupsOfCoffee", "hoursSleep", "hoursInSocialMedia", "hoursOnResearch", "minsInNovelReading", "workout")

```

Column {data-width=450}
-----------------------------------------------------------------------

### Average day pie chart

```{r}
colNames = c("cupsOfCoffee", "hoursSleep", "hoursInSocialMedia", "hoursOnResearch", "minsInNovelReading", "workout", "otherTasks")
meanCoffee = mean(newSelfDf$cupsOfCoffee)
meanSleep = mean(newSelfDf$hoursSleep)
meanMedia = mean(newSelfDf$hoursInSocialMedia)
meanResearch = mean(newSelfDf$hoursOnResearch)
meanNovel = mean(newSelfDf$minsInNovelReading) / 60.0
meanWorkout = mean(newSelfDf$workout)

totalMeans = meanCoffee + meanSleep + meanMedia + meanResearch + meanNovel + meanWorkout
meanOtherTasks = 24 - totalMeans

valuesForPieChart = c(meanCoffee, meanSleep, meanMedia, meanResearch, meanNovel, meanWorkout, meanOtherTasks)
plot_ly(newSelfDf, labels = colNames, values = valuesForPieChart, type = "pie") %>%
  layout(title = "What's my day like?")
```

Column {data-width=350}
-----------------------------------------------------------------------

### Analyzing how much sleep I get depending on cups of coffee I drink in a day

```{r}
scatterCoffeeSleep <- ggplot(newSelfDf, aes(x = cupsOfCoffee, y = hoursSleep, group = cupsOfCoffee )) + geom_boxplot()
scatterCoffeeSleep
ggplotly(scatterCoffeeSleep)
```

Column {data-width=450}
-----------------------------------------------------------------------

###Mostly my time of 9 hours is spent of research and 1 hour of novel reading.Reading is mostly in between 30mins to 1 hour while mostly research is 6-9 hours. Negative correlation is seen, hoursOnResearch goes down and minsInNovelReading goes up.


```{r}
heatResearchNovel <- ggplot(newSelfDf, aes(x = minsInNovelReading, y = hoursOnResearch)) + stat_binhex() + scale_x_continuous(breaks = seq(0, 120, by = 15)) + scale_fill_gradient(name = "Frequency")
heatResearchNovel
```

### Hours spent on research on a daily basis. So after looking at plot, I study for 5.5 hours 

```{r}
research <- ggplot(newSelfDf, aes(hoursOnResearch)) + labs(y = "Number of Days") + ggtitle("Hours Spent on Research")
research + geom_bar() + scale_x_continuous(breaks = seq(3, 9, by = 1))
```

